---
title: "H1N1 Exploration"
author: "Jose Figueroa, Liam Donohoe"
date: "April 6, 2020"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    df_print: kable
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
    df_print: kable
---

# Introduction

The goal of this project is to create a visualization of the progression of past epidemic or pandemic viruses throughout their lifetimes. In doing this, we hope to be able to provide a better understanding of not only the rate at which these types of viruses can spread throughout the population, but also the time it takes for scientists to develop and distribute a vaccine or cure for the spreading virus.

# Dataset Exploration
```{r, echo=FALSE, results=FALSE}
library(dplyr)
library(geojsonio)
library(here)
library(leaflet)
library(sp)
```


Dimensions:
```{r}
H1N1.df <- here("Data", "Pandemic (H1N1) 2009.csv") %>%
    read.csv(encoding = "UTF-8", stringsAsFactors = FALSE)
#H1N1.df$Country <- as.character(H1N1.df$Country)
Grand.total.df <- H1N1.df %>% filter(Country == "Grand Total")
H1N1.df <- H1N1.df %>% filter(Country != "Grand Total")
```
Dataframe Summary:
```{r}
glimpse(H1N1.df)
summary(H1N1.df)
```


Head:
```{r}
head(H1N1.df)
```
Tail: 
```{r}
tail(H1N1.df)
```
Unique Dates:
```{r}
unique.dates <- levels(H1N1.df$Update.Time)
length(unique.dates)
unique.dates
```

There are 22 unique weeks to account for. We should look into how to plot on a world map style heatmap (I have an idea on this), and then allowing for the option to traverse time on said heatmap. We can also easily transorm this notebook into a webapp with [RShiny](https://shiny.rstudio.com/)

# Single Date Plot

For this week, we can probably focus on having just 1 week plotted (5/23/2009). Then we can generalize to be interactive through time

```{r}
country.shapes <- here("Data", "countries.geojson") %>%
  geojson_read(what = "sp")
```


```{r}
CASES <- as.data.frame(country.shapes$ADMIN)
colnames(CASES) <- c("Country")
rownames(CASES) <- CASES$Country
CASES$Cases <- 0
CASES[H1N1.df[H1N1.df$Update.Time == unique.dates[1],"Country"], "Cases"] <- 
  H1N1.df[H1N1.df$Update.Time == unique.dates[1],"Cases"]
pal <- colorNumeric(
  palette = "YlOrRd",
  domain = country.shapes$CASES
)
country.shapes$CASES <- CASES$Cases
```


## May 23 2009

```{r}
m <- country.shapes %>% 
  leaflet()  %>%
  addPolygons(
    stroke=FALSE,
    smoothFactor = 0.2,
    fillOpacity = 0.7,
    color = ~pal(CASES)
  )%>%
  setView(-96.8, 39, 4)
m
```

```{r}
m2 <- country.shapes %>%
  leaflet(options=leafletOptions(minZoom=2, maxZoom=18)) %>%
  setView(-96.8, 39, 4) %>%
  addTiles(options = tileOptions(noWrap = TRUE)) %>%
  #addProviderTiles("MapBox", options = providerTileOptions(id = "mapbox.light")) %>%
  addPolygons(
    stroke=FALSE,
    smoothFactor = 0.2,
    fillOpacity = 0.7,
    color = ~pal(CASES))
m2
```

## "6/1/09 6:00"

```{r}
CASES <- as.data.frame(country.shapes$ADMIN)
colnames(CASES) <- c("Country")
rownames(CASES) <- CASES$Country
CASES$Cases <- 0
CASES[H1N1.df[H1N1.df$Update.Time == unique.dates[6],"Country"], "Cases"] <- 
  H1N1.df[H1N1.df$Update.Time == unique.dates[6],"Cases"]
pal <- colorNumeric(
  palette = "YlOrRd",
  domain = country.shapes$CASES
)
country.shapes$CASES <- CASES$Cases
```

```{r}
m2 <- country.shapes %>%
  leaflet(options=leafletOptions(minZoom=2, maxZoom=18)) %>%
  setView(-96.8, 39, 4) %>%
  addTiles(options = tileOptions(noWrap = TRUE)) %>%
  #addProviderTiles("MapBox", options = providerTileOptions(id = "mapbox.light")) %>%
  addPolygons(
    stroke=FALSE,
    smoothFactor = 0.2,
    fillOpacity = 0.7,
    color = ~pal(CASES))
m2
```

## "6/10/09 6:00"

```{r}
CASES <- as.data.frame(country.shapes$ADMIN)
colnames(CASES) <- c("Country")
rownames(CASES) <- CASES$Country
CASES$Cases <- 0
CASES[H1N1.df[H1N1.df$Update.Time == unique.dates[7],"Country"], "Cases"] <- 
  H1N1.df[H1N1.df$Update.Time == unique.dates[7],"Cases"]
pal <- colorNumeric(
  palette = "YlOrRd",
  domain = country.shapes$CASES
)
country.shapes$CASES <- CASES$Cases
```

```{r}
m2 <- country.shapes %>%
  leaflet(options=leafletOptions(minZoom=2, maxZoom=18)) %>%
  setView(-96.8, 39, 4) %>%
  addTiles(options = tileOptions(noWrap = TRUE)) %>%
  #addProviderTiles("MapBox", options = providerTileOptions(id = "mapbox.light")) %>%
  addPolygons(
    stroke=FALSE,
    smoothFactor = 0.2,
    fillOpacity = 0.7,
    color = ~pal(CASES))
m2
```

## "6/17/09 12:00"

```{r}
CASES <- as.data.frame(country.shapes$ADMIN)
colnames(CASES) <- c("Country")
rownames(CASES) <- CASES$Country
CASES$Cases <- 0
CASES[H1N1.df[H1N1.df$Update.Time == unique.dates[11],"Country"], "Cases"] <- 
  H1N1.df[H1N1.df$Update.Time == unique.dates[11],"Cases"]
pal <- colorNumeric(
  palette = "YlOrRd",
  domain = country.shapes$CASES
)
country.shapes$CASES <- CASES$Cases
```

```{r}
m2 <- country.shapes %>%
  leaflet(options=leafletOptions(minZoom=0, maxZoom=18)) %>%
  setView(-96.8, 39, 4) %>%
  addTiles(options = tileOptions(noWrap = TRUE)) %>%
  #addProviderTiles("MapBox", options = providerTileOptions(id = "mapbox.light")) %>%
  addPolygons(
    stroke=FALSE,
    smoothFactor = 0.2,
    fillOpacity = 0.7,
    color = ~pal(CASES))
m2
```